---
title: "shape_based_oligos"
format: html
editor: source
---

```{r}
#| echo: false
library(tidyr)
library(dplyr)
library(stringr)
library(seqinr)
library(R4RNA)
```


Parsing Dot Bracket into a more useful format for me - i,j pairing - i = 1-300 - j = i or 0, if unpaired

. -\> j = 0

sub table: bracket type, index (, 10 (, 11 (, 12 ), 15 ), 16

-\> look at https://stackoverflow.com/questions/29991917/indices-of-matching-parentheses-in-python https://stackoverflow.com/questions/61961836/find-and-label-matching-pairs-of-parentheses \# i, j dictionary -\> 2 contincous runs i and j = 1 ds, then sort into stems \# def stem as completed set of brackets \# all combos of helixes wi a stem --\> calling what a helix is will be complex \# limit n helicies per stem, divide helices by length of interupting runs on each side, like if 1-2 on one side but 5 on other, keep defining as same helix
Functions needed:

```{r used-functions}
#| echo: false
# turns a dot-bracket represenation of a structure into a table with with each i and j coordinate only appearing once
db_to_df_one_to_one <- function(dot_bracket_structure) {
  db <- dot_bracket_structure |> # takes string of dots and brackets
    str_split("") |>
    unlist() # and turns into a vector of characters
  len <- length(db) # length of sequence represented by dot bracket
  df <- tibble( # initialize tibble to store each base (i) and what it pairs to (j)
    i = 1:len,
    j = NA,
    value = NA) # temporary, used to track the order of open brackets and if they have been paired to their close bracket yet
    counter <- 1 # initialize counter, assigned to "value" when an open bracket is encountered
   #====================== iteration through the dot bracket ==============
    for (idx in 1:len){ # for each index in a dot bracket representation
      paired <- db[idx] # value at index
      if (paired == "."){
        df$j[idx] <- 0 # set j to 0; ie, not paired with anything
      } else if (paired == "(") {
        df$value[idx] <- counter 
        counter <- counter + 1 # increment counter
      } else if (paired == ")") {
        pair_bracket_index <- which.max(df$value) # index of highest value aka most recent opening bracket
        df$j[idx] <- df$i[pair_bracket_index] # assign index of most recent open bracket to j for this i (index x)
        df$value[pair_bracket_index] <- 0 # set the value of the base just paired to i to 0 so it can't be assigned to another i
      } else {
        print("Error! Unexpected character in dot bracket notation")
      }
    }
    df_out <- df |>
      select(!value) |># drop value column since it doesn't mean anything biologically
      drop_na() # drop rows with j = NA, ie open bracket, to represent each bp only once in a df
    return(df_out)
}  
# numbers helices
call_helicies <- function(i_j_pair_df) {
  i_j_pair_df <- i_j_pair_df |>
    mutate(runj = (c(0,diff(j))),  # need to pad with a zero to keep length
         runi = (c(0,diff(i)))) |> # run columns show the difference between the value in the row (i N or j N) and that in the previous (i N-1 or j N-1)
    mutate(in_helix = case_when( 
      j != 0 & i != 0 ~ 1, # if there are both coordinates, then in helix
      .default = 0
    ),
    helix_num = 0)
  helix_counter <- 0
  # now labeling by helix, defined as runs of base pairs where both i and j change by 1 from the previous pair
  for (idx in 1:nrow(i_j_pair_df)){
    if (i_j_pair_df$in_helix[idx] == 0){ #if not in helix, don't do anything
      next 
    } else if (i_j_pair_df$in_helix[idx] == 1 & i_j_pair_df$in_helix[idx-1] == 0) { # typical helix break where helicies are sep by some unpaired regions
      helix_counter <- helix_counter + 1
      i_j_pair_df$helix_num[idx] <- helix_counter
    } else if (i_j_pair_df$in_helix[idx] == 1 & i_j_pair_df$runj[idx] < -1) { # atypical helix start, when j jumps
      helix_counter <- helix_counter + 1
      i_j_pair_df$helix_num[idx] <- helix_counter
    } else if (i_j_pair_df$in_helix[idx] == 1 & i_j_pair_df$in_helix[idx-1] != 0) {
      i_j_pair_df$helix_num[idx] <- helix_counter
    } else {
      i_j_pair_df$helix_num[idx] <- "error"
    }
  }
  return(i_j_pair_df)
}
# numbers runs of helices/ stems
call_stems <- function(helix_df, n = 2) { # n = number of unpaired nt to ignore when itneruppting a step
  helix_rle <- rle(helix_df$in_helix)
  
  helix_rle_df <- tibble(
    lengths = helix_rle$lengths,
    values = helix_rle$values,
    in_run = NA
  )
  # decompress rle, but ignore lengths < n if value == 0 and call as part of run
  ## run calling
  for (idx in seq_along(helix_rle_df$lengths)) {
    if (idx == 1) { # for the first row
      helix_rle_df$in_run[idx] <- helix_rle_df$values[idx] # call in run or not by if in helix or not
    } else if (helix_rle_df$values[idx]) { # if in a helix
      helix_rle_df$in_run[idx] <- 1 # set value to 1/ in run
    } else if (!helix_rle_df$values[idx]) { # if this row is not a helix
      if (helix_rle_df$lengths[idx] > n) { # if more than n nt to ignore interrupting a run of helicies
        helix_rle_df$in_run[idx] <- 0 # break in run
      } else {
        helix_rle_df$in_run[idx] <- 1 # if lengths <= n, keep run going
      }
    } else {
       print("error!")
      }
  }
  ## run numbering  
  helix_rle_df <- helix_rle_df |>
      mutate(run_number = 0)
  run_idx <- 0
  previous_state <- 0 # if the previous row was in or out of a run
  for (idx in seq_along(helix_rle_df$lengths)) {
    if (helix_rle_df$in_run[idx] & !previous_state) { # if  starting a run
      run_idx <- run_idx + 1 # incriment counter
      helix_rle_df$run_number[idx] <- run_idx # save value in column
      previous_state <- helix_rle_df$in_run[idx] # save last state of in/out of a run to reference next loop
    } else if (helix_rle_df$in_run[idx] & previous_state){ # if continuing a run
      helix_rle_df$run_number[idx] <- run_idx
      previous_state <- helix_rle_df$in_run[idx]
    } else
      previous_state <- helix_rle_df$in_run[idx]
  }
  ## decompress rle, add in_run and run_num to y/ called helicies
  # https://devtut.github.io/r/run-length-encoding.html#identifying-and-grouping-by-runs-in-base-r
  in_runs_decomp <- rep(helix_rle_df$in_run, helix_rle_df$lengths)
  run_num_decomp <- rep(helix_rle_df$run_number, helix_rle_df$lengths)
  # bring everything together
  helix_df <- helix_df |>
    mutate(in_stem = in_runs_decomp,
           stem_num = run_num_decomp)
  return(helix_df)
}
# generates a complementary base
get_complement <- function(base) {
  comp_base <- switch(EXPR = base, # https://www.geeksforgeeks.org/switch-case-in-r/ example 2
           "A" = "U",
           "U" = "A",
           "C" = "G",
           "G" = "C")
  return(comp_base)
}
# mutates a single base
mutate_base <- function(base) {
   mutant_bases <- switch(EXPR = base, # https://www.geeksforgeeks.org/switch-case-in-r/ example 2
           "A" = c("U", "C", "G"),
           "U" = c("A", "C", "G"),
           "C" = c("U", "A", "G"),
           "G" = c("U", "A", "C"))
   mutant <- sample(mutant_bases, 1)
  return(mutant)
}

mutate_bases_by_idx <- function(idx_list, working_oligo, le_seq, valid_seq = 21:280) {
      for (idx in idx_list) { # for each index in a list
        if (idx %in% valid_seq) { # if index is not in primer regions
          substring(x = working_oligo, first = idx, last = idx) <- mutate_base(substr(x = le_seq, start = idx, stop = idx)) # mutate it
          }
      }
  return(working_oligo) # return mutated oligo
}

```

```{r, eval=FALSE, reference-functions}
### reference function!
#| ## from scramble-LE-MPRA.Rmd
# mutates a DNA string such that none of the bases have their original identity
mutate_dna <- function(string) {
  new_seq <- c()
  list_bases <- string |>
    strsplit("") |> #make into a list of characters
    unlist()
  for (i in 1:length(list_bases)){ # switch statement to turn a base into any base that is not the orginal, more compact than many if statements
    mutant_bases <- switch(EXPR = list_bases[i], # https://www.geeksforgeeks.org/switch-case-in-r/ example 2
           "A" = c("T", "C", "G"),
           "T" = c("A", "C", "G"),
           "C" = c("T", "A", "G"),
           "G" = c("T", "A", "C"))
   new_seq[i] <- sample(mutant_bases, 1)
  }
  output <- paste(new_seq, collapse = "")
  return(output)
}
# makes mutations based on predicted structures
mutate_structure <- function(LEsequence, viennaDotBracket, LEname, n_times = 3) { ##! add n_times for loop outside everything else
  # initialize lists to store seqs and names
  seqs <- c()
  names <- c()
  # use R4RNA to convet the dot braket to a table of base pairs by index
  basepairs <- viennaDotBracket %>% # data frame of pairs (i and j) by position
    viennaToHelix(.)
  basepairs_collapse <- basepairs %>% # start of each helix, arranged by index of the the "left" strand
    collapseHelix(.) %>%
    arrange(., i)
  # numbering helices
  helix_number <- 0
  row_num <- 1
  len_bp <- nrow(basepairs) #used to stop an error with while statement
  for (x in 1:nrow(basepairs_collapse)) {
    h_end <- basepairs_collapse$i[x] + basepairs_collapse$length[x] - 1 #calc helix end
    helix_number <- helix_number + 1
    while (basepairs$i[row_num] <=  h_end) { #while the position index is <= helix end index
        basepairs$value[row_num] <- helix_number
        row_num <- row_num + 1
      if (row_num > len_bp){ #deals with an error when row_num > number of rows in the df
        break
      }
    }
  }
  # left mutations (to the 'i' indices)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$i) # start of helix
    e <- max(helix_x$i) # end of helix
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    downstream <- str_sub(LEsequence, start = e+1, end = 260) #sequence downstream of window
    for (y in 1:n_times) {
      mutated <- str_sub(LEsequence, start = s, end = e) %>%
        mutate_dna()
      seqs[length(seqs)+1] <- paste(upstream, mutated, downstream, sep = "") %>% #append sequence to list
        paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
      names[length(names)+1] <- LEname %>%
        paste(., "_", as.character(s), ":", as.character(e), "_", as.character(y), #based on "n_times variable"
              sep = "")
    }
  }
  
  # right mutations (to the 'j' indices)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$j)
    e <- max(helix_x$j)
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    downstream <- str_sub(LEsequence, start = e+1, end = 260) #sequence downstream of window
      for (y in 1:n_times) {
        mutated <- str_sub(LEsequence, start = s, end = e) %>%
          mutate_dna()
        seqs[length(seqs)+1] <- paste(upstream, mutated, downstream, sep = "") %>% #append sequence to list
          paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
        names[length(names)+1] <- LEname %>%
          paste(., "_", as.character(s), ":", as.character(e), "_", as.character(y), #based on "n_times variable"
                sep = "")
    }
  }
  # compensatory (to both)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$i)
    e <- max(helix_x$i)
    s_j <- min(helix_x$j)
    e_j <- max(helix_x$j)
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    midstream <- str_sub(LEsequence, start = e+1, end = s_j-1) #sequence between helicies
    downstream <- str_sub(LEsequence, start = e_j+1, end = 260)
    for (y in 1:n_times) {
      mutated <- str_sub(LEsequence, start = s, end = e) %>%
        mutate_dna()
      list_bases <- mutated %>%
        strsplit(., "") %>% #make into a list of characters
        unlist(.)
      complement <- c()
      for (i in 1:length(list_bases)){ #getting comlementry seq
        complement[i] <- switch(EXPR = list_bases[i], 
             "A" = "T",
             "T" = "A",
             "C" = "G",
             "G" = "C")
      }
      rev_comp <- complement %>%
        rev(.) %>%
        paste(., collapse = "")
      seqs[length(seqs)+1] <- paste(upstream, mutated, midstream, rev_comp, downstream, sep = "") %>% #append sequence to list # complement should be rev_comp?
        paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
      names[length(names)+1] <- LEname %>%
        paste(., "_", as.character(s), ":", as.character(e),"::", 
        as.character(s_j), ":", as.character(e_j), "_", as.character(y), #based on "n_times variable"
              sep = "")
    }
  }
  output <- cbind(names, seqs) %>%
    as.data.frame(.)
  return(output)
}
single_stem_mutants <- function(stem_df, le_seq, le_name, n_rep = 3) {
  oligo_list <- c() # initialize lists to to store info
  names_list <- c()
  # get number of stems
  n_stem <- stem_df |>
    filter(in_stem == 1) |>
    pull(stem_num) |>
    max()
  # get i and j for each stem
  for (stem in 1:n_stem) { # for each stem
    i_idx <- stem_df |> # all i indicies in a stem
      filter(stem_num == stem, i !=0) |>
      pull(i)
    j_idx <- stem_df |> # all j indicies in a stem
      filter(stem_num == stem, j !=0) |>
      pull(j)
    ## making mutants
    rep_count <- 0
    while (rep_count < n_rep) {
      rep_count <- rep_count + 1
      wo <- le_seq
      oligo_list <- c(oligo_list, mutate_bases_by_idx(working_oligo = wo, le_seq = le_seq, idx_list = i_idx))
      names_list <- c(names_list, paste0(le_name,"_stem", as.character(stem), "i_rep", as.character(rep_count)))
    }
  }
  output <- tibble(
    names = names_list,
    oligos = oligo_list
  )
  return(output)
}
```

```{r}
t2 <- R2easyR::read.ct("LE_SHAPE_Summary_2/CT_files/Trak2_LE_structures.ct")
#t2_t <- R2easyR::add.dot.bracket(t2)
# throwing error: Warning: NAs introduced by coercionError in if (as.numeric(as.character(ctdata_frame$BP[i])) > as.numeric(as.character(ctdata_frame$N[i]))) { : missing value where TRUE/FALSE needed, so just using https://rna.urmc.rochester.edu/RNAstructureWeb/Servers/ct2dot/ct2dot.html
t2sq <- "GCUUCGAUAUCCGCAUGCUAAACCACUAACUGACCUCGUGUAUAAAUUAGCACAUGAAGGAGAGAGAUGCACUGAGAGUGGAGACCGGCCUAACUUGAGAGAAAAGGGAUGUUGCAGAAGGACUGUGAAUGUGGAGAGAGCGCAGUGAAGGGAUAGGACGUUUGUUUGCAAGAGGGUCUUUGAGGAUGGAGUCUAGUAAACAAAGUGGGGACAAGUGAUGACCAGUUUGGAGACAAAUGGAAGAAAGGGUUAGUUUGCUCCUUCAGCUACGUUUCCUUGCCUCUUGCGGUCGCACUAGUG"

t2db <- ".........(((...(((....(((((..((..((((((((..........))))).)))..))............))))).(((...(((.............)))...))))))...)))...............((.(((((.((..((..(((((((((((((.((((..((((............((((..............)))).......))))..)))).)))))))))((((...((((......)))))))).........))))..))))))))).))........."

t2df <- t2db |>
  viennaToHelix() # i have no idea how helix represents dot bracket, lots of missing base pairs :/
```


```{r, echo=FALSE, eval=FALSE}
db_to_df <- function(dot_bracket_structure) {
  db <- dot_bracket_structure |> # takes string of dots and brackets
    str_split("") |>
    unlist() # and turns into a vector of characters
  
  len <- length(db) # length of sequence represented by dot bracket
  
  df <- tibble( # initialize tibble to store each base (i) and what it pairs to (j)
    i = 1:len,
    j = NA,
    value = NA) # temporary, used to track the order of open brackets and if they have been paired to their close bracket yet
    
    counter <- 1 # initialize counter, assigned to "value" when an open bracket is encountered
   #====================== iteration through the dot bracket ==============
     
    for (idx in 1:len){ # for each index in a dot bracket representation
      paired <- db[idx] # value at index
      if (paired == "."){
        df$j[idx] <- 0 # set j to 0; ie, not paired with anything
      } else if (paired == "(") {
        df$value[idx] <- counter 
        counter <- counter + 1 # increment counter
      } else if (paired == ")") {
        pair_bracket_index <- which.max(df$value) # index of highest value aka most recent opening bracket
        df$j[idx] <- df$i[pair_bracket_index] # assign index of most recent open bracket to j for this i (index x)
        df$value[pair_bracket_index] <- 0 # set the value of the base just paired to i to 0 so it can't be assigned to another i
      } else {
        print("Error! Unexpected character in dot bracket notation")
      }
    }
   #============== backfilling j values for indicies of open brackets =========
    for (idx in 1:len){
      val <- df$j[idx] 
      if (is.na(val)) { # if the value in j at row idx is NA (indicate open bracket)
       other_half <- which(df$j == idx) # index of row where idx = j
       df$j[idx] <- other_half # assign index of matching close bracket to j
      }
    }
    df_out <- df |>
      select(!value) # drop value column since it doesn't mean anything biologically
    return(df_out)
}

```

```{r}
# https://stackoverflow.com/questions/16118050/how-to-check-if-a-vector-contains-n-consecutive-numbers
# use diff() to find runs of numbers where the difference between them == 1
# then find runs of at least 2, tolerating some number of runs between
x <- db_to_df_one_to_one(t2db)

xt <- x |>
  mutate(runj = (c(0,diff(j))),  #need to pad with a zero to keep length
         runi = (c(0,diff(i)))) # run columns show the difference between the value in the row (i N or j N) and that in the previous (i N-1 or j N-1)
helixn <- 0
xt <- xt |>
  mutate(helix = NA)
for (idx in 1:nrow(xt)){
  if (xt$runj[idx] == 0 | xt$j[idx] == 0){ # if no difference between js
    next # skip to the next row
  } else if (xt$runi[idx] == 1) { # if a helix continues
    xt$helix[idx] <- helixn
  } else if (xt$runj[idx] > 0){ # if there is a positive differnce in js, ie, a helix starts
    helixn <- helixn + 1 # increment helix number
    xt$helix[idx] <- helixn 
  } else if (xt$runj[idx] < -1) { # if helix ends
    next 
  } else {
    xt$helix[idx] <- "Error"
  }
}

```

```{r}
# approached by first labeling if is or is not a helix (i and j =! 0), then outing runs
x <- db_to_df_one_to_one(t2db)
y <- call_helicies(x)
# call stems by if i or j is continous, ignore bulges N or smaller (?)
n <- 2 # number of nt to ignore
  # 1. call runs of in helix vs not, with tolerance of skipping N nt
  #using RLE ->
helix_run_rle <- rle(y$in_helix)

helix_run_rle_df <- tibble(
  lengths = helix_run_rle$lengths,
  values = helix_run_rle$values,
  in_run = NA
)
# decompress rle, but ignore lengths < n if value == 0 and call as part of run
## run calling
for (idx in seq_along(helix_run_rle_df$lengths)) {
  if (idx == 1) { # for the first row
    helix_run_rle_df$in_run[idx] <- helix_run_rle_df$values[idx] # call in run or not by if in helix or not
  } else if (helix_run_rle_df$values[idx]) { # if in a helix
    helix_run_rle_df$in_run[idx] <- 1 # set value to 1/ in run
  } else if (!helix_run_rle_df$values[idx]) { # if this row is not a helix
    if (helix_run_rle_df$lengths[idx] > n) { # if more than n nt to ignore interrupting a run of helicies
      helix_run_rle_df$in_run[idx] <- 0 # break in run
    } else {
      helix_run_rle_df$in_run[idx] <- 1 # if lengths <= n, keep run going
    }
  } else {
     print("error!")
    }
}
## run numbering  
helix_run_rle_df <- helix_run_rle_df |>
  mutate(run_number = 0)
run_idx <- 0
prevous_state <- 0
for (idx in seq_along(helix_run_rle_df$lengths)) {
  if (helix_run_rle_df$in_run[idx] & !prevous_state) { # if  starting a run
    run_idx <- run_idx + 1 # incriment counter
    helix_run_rle_df$run_number[idx] <- run_idx # save value in column
    prevous_state <- helix_run_rle_df$in_run[idx] # save last state of in/out of a run to reference next loop
  } else if (helix_run_rle_df$in_run[idx] & prevous_state){ # if continuing a run
    helix_run_rle_df$run_number[idx] <- run_idx
    prevous_state <- helix_run_rle_df$in_run[idx]
  } else
    prevous_state <- helix_run_rle_df$in_run[idx]
}
## decompress rle, add in_run and run_num to y/ called helicies
# https://devtut.github.io/r/run-length-encoding.html#identifying-and-grouping-by-runs-in-base-r
in_runs_decomp <- rep(helix_run_rle_df$in_run, helix_run_rle_df$lengths)
run_num_decomp <- rep(helix_run_rle_df$run_number, helix_run_rle_df$lengths)
# bring everything together
y <- y |>
  mutate(in_run = in_runs_decomp,
         run_num = run_num_decomp)


z <- call_stems(y, n = 4)
```

next: actually making mutations
for each stem:
collect all i j pairs, exclude an where j = 0, from a stem N
mutate all i, then all j, then compensate (mutate i, then generate matching j sequence)
```{r}
# starting from z df and t2seq
# if i or j are in primer regions, just mutate once --> if i/j %in% primer; continue
z_stems <- z |>
  filter(in_stem == 1)
validseq <- 21:280
le_name <- "test"
n_stem <- 1
oligo_list <- c()
names_list <- c()

#lists of indices to mutate
i_1 <- z_stems |> # all i indicies for stem 1
  filter(stem_num == n_stem, i != 0) |>
  pull(i)
# https://stackoverflow.com/questions/36596928/replace-character-at-certain-location-within-string to edit oligo
working_oligo <- t2sq
for (idx in i_1) {
  substring(x = working_oligo, first = idx, last = idx) <- mutate_base(substr(x = t2sq, start = idx, stop = idx))
}
## returns working oligo, where mutations have been made

## naming
#lename_stemN_i_repN; lename_stemN_j_repN; lename_stemN_ij_repN
```

```{r making-mutations}
wo <- t2sq
a <- mutate_bases_by_idx(idx_list = i_1, working_oligo = wo, le_seq = wo)

```

```{r combinaitons}
oligo_list <- c() # initialize lists to to store info
names_list <- c()
# get number of stems
stem_df <- z
n_stem <- stem_df |>
  filter(in_stem == 1) |>
  pull(stem_num) |>
  max()
combos <- c()
for(n in 1:n_stem){
  temp_comb <- combn(x = 1:n_stem, m = n, simplify = FALSE)
  combos <- c(combos, temp_comb)
}
combos <- combos |>
    unname()
for (c in combos) { # for each combination of stems
    stems <- c |> unlist()
    i_idx <- stem_df |> # all i indicies in a 
      filter(stem_num %in% stems, i !=0) |>
      pull(i)
    j_idx <- stem_df |> # all j indicies in a 
      filter(stem_num %in% stems, j !=0) |>
      pull(j)
}
valid_seq <- 21:280
stems <- c(1,3)## compensatory testing
ij_df <- stem_df |> ## for compensatory, would go into for c in combos loop
    filter(stem_num %in% stems) |>
    select(i,j)
curr_oligo <- t2sq
for (idx in 1:nrow(ij_df)){
  if (ij_df$i[idx] %in% valid_seq & ij_df$j[idx] %in% valid_seq) { # if index is not in primer regions
    mutant <-  mutate_base(substr(x = t2sq, start = ij_df$i[idx], stop = ij_df$i[idx]))
    substring(x = curr_oligo, first = ij_df$i[idx], last = ij_df$i[idx]) <- mutant # mutate it add mutant to i index of oligo being built
    substring(x = curr_oligo, first = ij_df$j[idx], last = ij_df$j[idx]) <- get_complement(mutant) # put complement of i base in j position
    
          }  }

###########
combo_stem_mutants <- function(stem_df, le_seq, le_name, n_rep = 3) {
  oligo_list <- c() # initialize lists to to store info
  names_list <- c()
  # get number of stems
  n_stem <- stem_df |>
    filter(in_stem == 1) |>
    pull(stem_num) |>
    max()
  combos <- c()
  for(n in 1:n_stem){
  temp_comb <- combn(x = 1:n_stem, m = n, simplify = FALSE) # all combos for this m
  combos <- c(combos, temp_comb)
  }
  combos <- combos |>
    unname()
  # get i and j for each combo
  #######
  for (c in combos) { # for each combination of stems
    stems <- c |> unlist()
    i_idx <- stem_df |> # all i indicies in a 
      filter(stem_num %in% stems, i !=0) |>
      pull(i)
    j_idx <- stem_df |> # all j indicies in a 
      filter(stem_num %in% stems, j !=0) |>
      pull(j)
    ## making mutants
    rep_count <- 0
    while (rep_count < n_rep) {
      rep_count <- rep_count + 1
      oligo_list <- c(oligo_list, mutate_bases_by_idx(working_oligo = le_seq, le_seq = le_seq, idx_list = i_idx)) # mutating i
      names_list <- c(names_list, paste0(le_name,"_stem", paste(c,collapse = "-"), "i_rep", as.character(rep_count))) # helicies mutated separated by dashes
      oligo_list <- c(oligo_list, mutate_bases_by_idx(working_oligo = le_seq, le_seq = le_seq, idx_list = j_idx)) # mutating j
      names_list <- c(names_list, paste0(le_name,"_stem", paste(c,collapse = "-"), "j_rep", as.character(rep_count)))
    }
  }
  output <- tibble(
    names = names_list,
    oligos = oligo_list
  )
  return(output)
}
combo_test <- combo_stem_mutants(stem_df = z, le_seq = t2sq, le_name = "test", n_rep = 3)

## compendastory mutations 
compensatory_mutate <- function(stem_df, working_oligo, valid_seq = 21:280, stem_idxes) {
  ij_df <- stem_df |>
    filter(stem_num %in% stem_idexs) |>
    select(i,j)
  for (idx in ij_df){
    
  }
  
 #########   
  for (idx in idx_list) { # for each index in a list
      if (idx %in% valid_seq) { # if index is not in primer regions
        substring(x = working_oligo, first = idx, last = idx) <- mutate_base(substr(x = t2sq, start = idx, stop = idx)) # mutate it
        }
    }
  return(working_oligo) # return mutated oligo
}

```
