---
title: "shape_based_oligos"
format: html
editor: visual
---

```{r}
#| echo: false
library(tidyr)
library(dplyr)
library(stringr)
library(seqinr)
library(R4RNA)
```

Functions needed:

```{r functions}
#| echo: false
#| ## from scramble-LE-MPRA.Rmd
# mutates a DNA string such that none of the bases have their original identity
mutate_dna <- function(string) {
  new_seq <- c()
  list_bases <- string |>
    strsplit("") |> #make into a list of characters
    unlist()
  for (i in 1:length(list_bases)){ # switch statement to turn a base into any base that is not the orginal, more compact than many if statements
    mutant_bases <- switch(EXPR = list_bases[i], # https://www.geeksforgeeks.org/switch-case-in-r/ example 2
           "A" = c("T", "C", "G"),
           "T" = c("A", "C", "G"),
           "C" = c("T", "A", "G"),
           "G" = c("T", "A", "C"))
   new_seq[i] <- sample(mutant_bases, 1)
  }
  output <- paste(new_seq, collapse = "")
  return(output)
}



# makes mutations based on predicted structures
mutate_structure <- function(LEsequence, viennaDotBracket, LEname, n_times = 3) { ##! add n_times for loop outside everything else
  # initialize lists to store seqs and names
  seqs <- c()
  names <- c()
  # use R4RNA to convet the dot braket to a table of base pairs by index
  basepairs <- viennaDotBracket %>% # data frame of pairs (i and j) by position
    viennaToHelix(.)
  basepairs_collapse <- basepairs %>% # start of each helix, arranged by index of the the "left" strand
    collapseHelix(.) %>%
    arrange(., i)
  # numbering helices
  helix_number <- 0
  row_num <- 1
  len_bp <- nrow(basepairs) #used to stop an error with while statement
  for (x in 1:nrow(basepairs_collapse)) {
    h_end <- basepairs_collapse$i[x] + basepairs_collapse$length[x] - 1 #calc helix end
    helix_number <- helix_number + 1
    while (basepairs$i[row_num] <=  h_end) { #while the position index is <= helix end index
        basepairs$value[row_num] <- helix_number
        row_num <- row_num + 1
      if (row_num > len_bp){ #deals with an error when row_num > number of rows in the df
        break
      }
    }
  }
  # left mutations (to the 'i' indices)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$i) # start of helix
    e <- max(helix_x$i) # end of helix
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    downstream <- str_sub(LEsequence, start = e+1, end = 260) #sequence downstream of window
    for (y in 1:n_times) {
      mutated <- str_sub(LEsequence, start = s, end = e) %>%
        mutate_dna()
      seqs[length(seqs)+1] <- paste(upstream, mutated, downstream, sep = "") %>% #append sequence to list
        paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
      names[length(names)+1] <- LEname %>%
        paste(., "_", as.character(s), ":", as.character(e), "_", as.character(y), #based on "n_times variable"
              sep = "")
    }
  }
  
  # right mutations (to the 'j' indices)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$j)
    e <- max(helix_x$j)
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    downstream <- str_sub(LEsequence, start = e+1, end = 260) #sequence downstream of window
      for (y in 1:n_times) {
        mutated <- str_sub(LEsequence, start = s, end = e) %>%
          mutate_dna()
        seqs[length(seqs)+1] <- paste(upstream, mutated, downstream, sep = "") %>% #append sequence to list
          paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
        names[length(names)+1] <- LEname %>%
          paste(., "_", as.character(s), ":", as.character(e), "_", as.character(y), #based on "n_times variable"
                sep = "")
    }
  }
  # compensatory (to both)
  for (x in 1:helix_number) {
    helix_x <- basepairs %>%
      filter(value == x)
    s <- min(helix_x$i)
    e <- max(helix_x$i)
    s_j <- min(helix_x$j)
    e_j <- max(helix_x$j)
    upstream <- str_sub(LEsequence, start = 1, end = s-1) #sequence upstream of window
    midstream <- str_sub(LEsequence, start = e+1, end = s_j-1) #sequence between helicies
    downstream <- str_sub(LEsequence, start = e_j+1, end = 260)
    for (y in 1:n_times) {
      mutated <- str_sub(LEsequence, start = s, end = e) %>%
        mutate_dna()
      list_bases <- mutated %>%
        strsplit(., "") %>% #make into a list of characters
        unlist(.)
      complement <- c()
      for (i in 1:length(list_bases)){ #getting comlementry seq
        complement[i] <- switch(EXPR = list_bases[i], 
             "A" = "T",
             "T" = "A",
             "C" = "G",
             "G" = "C")
      }
      rev_comp <- complement %>%
        rev(.) %>%
        paste(., collapse = "")
      seqs[length(seqs)+1] <- paste(upstream, mutated, midstream, rev_comp, downstream, sep = "") %>% #append sequence to list # complement should be rev_comp?
        paste("GCTTCGATATCCGCATGCTA", ., "CTCTTGCGGTCGCACTAGTG", sep = "") #add PCR handles
      names[length(names)+1] <- LEname %>%
        paste(., "_", as.character(s), ":", as.character(e),"::", 
        as.character(s_j), ":", as.character(e_j), "_", as.character(y), #based on "n_times variable"
              sep = "")
    }
  }
  output <- cbind(names, seqs) %>%
    as.data.frame(.)
  return(output)
}
```

```{r}
t2 <- R2easyR::read.ct("LE_SHAPE_Summary_2/CT_files/Trak2_LE_structures.ct")
#t2_t <- R2easyR::add.dot.bracket(t2)
# throwing error: Warning: NAs introduced by coercionError in if (as.numeric(as.character(ctdata_frame$BP[i])) > as.numeric(as.character(ctdata_frame$N[i]))) { : missing value where TRUE/FALSE needed, so just using https://rna.urmc.rochester.edu/RNAstructureWeb/Servers/ct2dot/ct2dot.html
t2sq <- "GCUUCGAUAUCCGCAUGCUAAACCACUAACUGACCUCGUGUAUAAAUUAGCACAUGAAGGAGAGAGAUGCACUGAGAGUGGAGACCGGCCUAACUUGAGAGAAAAGGGAUGUUGCAGAAGGACUGUGAAUGUGGAGAGAGCGCAGUGAAGGGAUAGGACGUUUGUUUGCAAGAGGGUCUUUGAGGAUGGAGUCUAGUAAACAAAGUGGGGACAAGUGAUGACCAGUUUGGAGACAAAUGGAAGAAAGGGUUAGUUUGCUCCUUCAGCUACGUUUCCUUGCCUCUUGCGGUCGCACUAGUG"

t2db <- ".........(((...(((....(((((..((..((((((((..........))))).)))..))............))))).(((...(((.............)))...))))))...)))...............((.(((((.((..((..(((((((((((((.((((..((((............((((..............)))).......))))..)))).)))))))))((((...((((......)))))))).........))))..))))))))).))........."

t2df <- t2db |>
  viennaToHelix() # i have no idea how helix represents dot bracket, lots of missing base pairs :/
```

Parsing Dot Bracket into a more useful format for me - i,j pairing - i = 1-300 - j = i or 0, if unpaired

. -\> j = 0

sub table: bracket type, index (, 10 (, 11 (, 12 ), 15 ), 16

-\> look at https://stackoverflow.com/questions/29991917/indices-of-matching-parentheses-in-python https://stackoverflow.com/questions/61961836/find-and-label-matching-pairs-of-parentheses \# i, j dictionary -\> 2 contincous runs i and j = 1 ds, then sort into stems \# def stem as compeletewd set of brackets \# all combos of helixes wi a stem --\> calling what a helix is will be complex \# limit n helicies per stem, divide helices by length of interupting runs on each side, like if 1-2 on one side but 5 on other, keep defining as same helix

```{r}
db <- t2db |>
  str_split("") |>
  unlist()

len <- 300
df <- tibble(
  i = 1:len,
  j = NA,
  value = NA
)
counter <- 1
for (x in 1:len){ # for each index in a dot bracket representation
  paired <- db[x] # value at index
  if (paired == "."){
    df$j[x] <- 0 # set j to 0; ie, not paired with anything
  } else if (paired == "(") {
    df$value[x] <- counter 
    counter <- counter + 1 # increment counter
  } else if (paired == ")") {
    pair_bracket_index <- which.max(df$value) # index of highest value aka most recent opening bracket
    df$j[x] <- df$i[pair_bracket_index] # assign index of most recent open bracket to j for this i (index x)
    df$value[pair_bracket_index] <- 0 # set the value of the base just paired to i to 0 so it can't be assigned to another i
  } else {
    print("Error! Unexpected character in dot bracket notation")
  }
}

# Next, need to back fill j for all open brackets; set as NA currently

```

? trim to 260 after mutating ect